Data Exploration

### 2019-01-23 src/data_exploration.Rmd
ID cells (xyzt) with >1 observation

vif_allo2_depth_sq <- read.csv("../output/bio/vif_allo2_depth.csv", header = TRUE)
vif_allo2_depth_sq$GVIF2Dfsq <- vif_allo2_depth_sq$GVIF..1..2.Df..^2
write.csv(vif_allo2_depth_sq, "../output/bio/vif_allo2_depth_sq", row.names = TRUE)
vif_allo2_depth_sq
```

without gear
```{r}
vif_allbuttemp_surfacegear <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbuttemp_surfacegear, "../output/bio/vif_allbuttemp_surfacegear.csv", row.names = TRUE)
vif_allbuttemp_surfacegear

vif_allo2_depthgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allo2_depthgear, "../output/bio/vif_allo2_depthgear.csv", row.names = TRUE)
vif_allo2_depthgear
```

>>>>

vif_allo2_depth_sq <- read.csv("../output/bio/vif_allo2_depth.csv", header = TRUE)
vif_allo2_depth_sq$GVIF2Dfsq <- vif_allo2_depth_sq$GVIF..1..2.Df..^2
write.csv(vif_allo2_depth_sq, "../output/bio/vif_allo2_depth_sq.csv", row.names = TRUE)
vif_allo2_depth_sq
```

without gear
```{r}
vif_allbuttemp_surfacegear <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbuttemp_surfacegear, "../output/bio/vif_allbuttemp_surfacegear.csv", row.names = TRUE)
vif_allbuttemp_surfacegear

vif_allo2_depthgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allo2_depthgear, "../output/bio/vif_allo2_depthgear.csv", row.names = TRUE)
vif_allo2_depthgear
```

```{r}
vif_test <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
vif_test
```

# check to see how many cells have >1 point per timeslice

this needs to be 3D so you need to create a cell_id_3d
```{r}
presab$cell_id_3d <- paste(presab$cell_id, presab$depthlayerno, sep="_") #create a new column that is a unique cell_id & depth ID. 
head(presab)
```
and now do one with a timeslice label
```{r}
presab$cell_id_xyzt <- paste(presab$cell_id_3d, presab$year, presab$month, sep="_") #create a new column that is a unique cell_id & depth ID. 
head(presab)
```

the data is in presab$total_cell_obs_xyt 
```{r}
unique_pointpercell <- unique(presab$total_cell_obs_xyzt)
unique_pointpercell
```
bugger... ok so lets do a frequency table
```{r}
celobsmore1 <- subset(presab, total_cell_obs_xyzt > 1)
pointpercell <- with(celobsmore1, table(cell_id_xyzt, total_cell_obs_xyzt))
write.csv(pointpercell, file = "../output/bio/pointpercell.csv")
nrow(pointpercell)
pointpercell
```

so there are 949 3d timesliced cells that contain more than one occurrence. 
- different sampling days
- different coordinates
- risk of duplicate entries (but only if two instituions inputted the same data with slightly different coordinates and different reserch program names - see cell3985_18_2000_7 below as example)

```{r}
cell3985_18_2000_7  <- subset(celobsmore1, cell_id_xyzt == "3985_18_2000_7")
cell3985_18_2000_7 
```
